---
title: "Default of Credit Card Clients"
author: "John Richard Wrenn"
date: "December 20, 2018"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

Dataset is from University of California in Irvine, California

Data is from a Tiawanese bank

30,000 observations

23 Predictors with 1 binary target variable

## Variables Used

- Credit (> 0)
- Age (> 0)
- Gender (M, F)
- Education (High School, Uni, Grad School, Other)
- Marital Status (Single, Married, Other)
- Current Payment Status (-1, 0, 1, ..., 8)
- Current Bill (>= 0)
- Previous Bill (>= 0)
- Default (Yes, No)

## Intro Graphs

```{r}
data = read.csv(file = "C:\\Users\\student\\Documents\\Applied Data Mining\\Final Project\\default of credit card clients.csv", header = TRUE)

# Remove this NULLs variable, is ID only
data$X<-NULL

# Replace names
names(data) <- c('credit', 'gender', 'education', 'maritalStatus', 'age', 'sepPaymentStatus', 'augPaymentStatus', 'julPaymentStatus', 'junPaymentStatus', 'mayPaymentStatus', 'aprPaymentStatus', 'sepBillStatement', 'augBillStatement', 'julBillStatement', 'junBillStatement', 'mayBillStatement', 'aprBillStatement', 'sepPrevBillStatement', 'augPrevBillStatement', 'julPrevBillStatement', 'junPrevBillStatement', 'mayPrevBillStatement', 'aprPrevBillStatement', 'target')

# Make odd values missing
data$education[data$education == 6]<-NA
data$education[data$education == 5]<-NA
data$education[data$education == 0]<-NA

data$maritalStatus[data$maritalStatus == 0]<-NA

data$sepPaymentStatus[data$sepPaymentStatus == -2]<-NA
data$augPaymentStatus[data$augPaymentStatus == -2]<-NA
data$julPaymentStatus[data$julPaymentStatus == -2]<-NA
data$junPaymentStatus[data$junPaymentStatus == -2]<-NA
data$mayPaymentStatus[data$mayPaymentStatus == -2]<-NA
data$aprPaymentStatus[data$aprPaymentStatus == -2]<-NA

data$sepBillStatement[data$sepBillStatement < 0]<-NA
data$augBillStatement[data$augBillStatement < 0]<-NA
data$julBillStatement[data$julBillStatement < 0]<-NA
data$junBillStatement[data$junBillStatement < 0]<-NA
data$mayBillStatement[data$mayBillStatement < 0]<-NA
data$aprBillStatement[data$aprBillStatement < 0]<-NA

data$sepPrevBillStatement[data$sepPrevBillStatement < 0]<-NA
data$augPrevBillStatement[data$augPrevBillStatement < 0]<-NA
data$julPrevBillStatement[data$julPrevBillStatement < 0]<-NA
data$junPrevBillStatement[data$junPrevBillStatement < 0]<-NA
data$mayPrevBillStatement[data$mayPrevBillStatement < 0]<-NA
data$aprPrevBillStatement[data$aprPrevBillStatement < 0]<-NA

# Make some ints factors
target.f <- factor(data$target, labels = c("No", "Yes"))
gender.f <- factor(data$gender, labels = c("M", "F"))
edu.f <- factor(data$education, labels = c("GradSchool", "University", "HighSchool", "Other"))
marital.f <- factor(data$maritalStatus, labels = c("Married", "Single", "Other"))

data$target <- target.f
data$gender <- gender.f
data$education <- edu.f
data$maritalStatus <- marital.f

library(ggplot2)

ggplot(data) + geom_bar(aes(data$target))
```


## Intro Graphs (cont.)

```{r}
ggplot(data) + geom_density(aes(data$credit))
```

## Intro Graphs (cont.)

```{r}
ggplot(data) + geom_bar(aes(data$sepPaymentStatus, color = data$gender), position = "dodge", na.rm = TRUE)
```

## Intro Graphs (cont.)

```{r, echo = FALSE}
data.male = data[data$gender == "M",]
data.female = data[data$gender == "F",]
data.female.under = data.female[sample(1:nrow(data.female), nrow(data.male)),]
data.gender.balance = rbind(data.male, data.female.under)
ggplot(data.gender.balance) + geom_bar(aes(data.gender.balance$sepPaymentStatus, color = data.gender.balance$gender), position = "dodge", na.rm = TRUE)
```

## Intro Graphs (cont.)

```{r, echo = FALSE}
ggplot(data) + geom_density(aes(data$credit, color = data$gender)) + facet_wrap(~target)
```

## Intro Graphs (cont.)

```{r, echo = FALSE}
ggplot(data) + geom_freqpoly(aes(data$credit, color = data$education), bins = 30)
```

## rpart Results

```{r}
rpart.list <- c("Model", "Median", "Knn", "Complete", "Bin Median", "Bin Knn", "Bin Complete", "Dummy", "Accuracy", "69.35", "67.74", "69.84", "69.45", "69.80", "68.42", "70.95", "sensitivity", "52.61", "50.95", "52.81", "56.88", "56.63", "51.35", "53.89", "Specificity", "86.08", "84.52", "87.95", "82.01", "82.96", "85.84", "89.10")
rpart.matrix <- matrix(rpart.list, nrow = length(rpart.list)/4, ncol = 4)
rpart.matrix
```

## C5.0 Results

```{r}
C5.0.list <- c("Model", "Median", "Knn", "Complete", "Bin Median", "Bin Knn", "Bin Complete", "Dummy", "Size", "26", "28", "13", "169", "177", "107", "11", "Error Rate", "28.50%", "27.40%", "26.30%", "24.70%", "25.60%", "25.10%", "26.80%")
C5.0.matrix <- matrix(C5.0.list, nrow = length(C5.0.list)/3, ncol = 3)
C5.0.matrix
```

## Ranger Results

```{r}
ranger.list <- c("Model", "Median", "Knn", "Complete", "Bin Median", "Bin Knn", "Bin Complete", "Dummy", "Accuracy", "70.87%", "70.84%", "72.65%", "70.86%", "70.79", "70.96%", "72.23", "Min Node Size", "3", "3", "2", "2", "3", "3", "3")
ranger.matrix <- matrix(ranger.list, nrow = length(ranger.list)/3, ncol = 3)
ranger.matrix
```
## rpart Tuning Results

```{r}
rpart.tune.list <- c("Model", "Median", "Knn", "Complete", "Bin Median", "Bin Knn", "Bin Complete", "Dummy", "Accuracy", "69.35", "67.74", "69.84", "69.45", "69.80", "68.42", "70.95", "sensitivity", "52.61", "50.95", "52.81", "56.88", "56.63", "51.35", "53.89", "Specificity", "86.08", "84.52", "87.95", "82.01", "82.96", "85.84", "89.10")
rpart.tune.matrix <- rpart.matrix
rpart.tune.matrix
```

## C5.0 Tuning Results

```{r}
C5.0.tune.list <- c("Model", "Median", "Knn", "Complete", "Bin Median", "Bin Knn", "Bin Complete", "Dummy", "Error Rate", "27.70%", "26.50%", "26.10%", "24.40%", "25.00%", "24.30%", "26.80%", "Max Tree Size", "26", "38", "16", "169", "117", "107", "11")
C5.0.tune.matrix <- matrix(C5.0.tune.list, nrow = length(C5.0.tune.list)/3, ncol = 3)
C5.0.tune.matrix
```

## Ranger Tuning Results

```{r}
ranger.tune.list <- c("Model", "Median", "Knn", "Complete", "Bin Median", "Bin Knn", "Bin Complete", "Dummy", "Accuracy", "70.67%", "71.08%", "72.58%", "70.89%", "71.01%", "71.24%", "72.34%", "Min Node Size", "2", "1", "1", "3" , "1", "3", "2", "Split Rule", "Gini", "Gini", "Extra Trees", "Gini", "Gini", "Gini", "Gini")
ranger.tune.matrix <- matrix(ranger.tune.list, ncol = 4, nrow = length(ranger.tune.list)/4)
ranger.tune.matrix
```

## Conclusion

The C5.0 algorithm is the superior algorithm in terms of error rate

Binning can generally improve accuracy

## Cool Fact and Feature

```{r, echo = FALSE}
sizes <- sapply(ls(), function(n) object.size(get(n)), simplify = FALSE)

#print(sapply(sizes[order(as.integer(sizes))], function(s) format(s, unit = 'auto')))
```

```{r, echo = FALSE}
sum <- 0
for(i in 1:length(sizes)) {
  sum = sum + as.integer(sizes[i])
}
```
sizes <- sapply(ls(), function(n) object.size(get(n)), simplify = FALSE)

print(sapply(sizes[order(as.integer(sizes))], function(s) format(s, unit = 'auto')))

720 MB